---
title: 'Data Products Week 3: North American Peaks'
author: "R. A. Reitmeyer"
date: "February 26, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

This is a simple presentation of North American Mountain Peaks
in a plotly interactive graph. It was created February 26, 2017 
for the week 3 project of the "Data Products" coursera course.

The purpose of the plot is to illustrate the relationship between
peak isloation (distance to nearest-higher peak) and prominance 
(height above lowest point connecting to nearest-higher peak).


## Data Source

Data comes from a wikipedia:
[https://en.wikipedia.org/wiki/Table_of_the_highest_major_summits_of_North_America]

Using the KML file would have been much easier, but the KML lacks the elevation,
prominance and isolation information.

## Data Processing

Data was extracted from the main table in the HTML page and parsed.

```{r}
library(XML)
library(stringr)

tree <- htmlParse('./en.wikipedia.org_wiki_Table_of_the_highest_major_summits_of_North_America.html')
data <- do.call(rbind, 
    xpathApply(tree, '/html/body//table[@class="wikitable sortable"]/tr',
        function(tr) {
            tds<-getNodeSet(tr, "./td")
            l <- length(tds)
            if (l > 7) { 
                data.frame(
                    raw_name=xmlValue(tds[[l-6]]), 
                    region=xmlValue(tds[[l-5]]), 
                    range=xmlValue(tds[[l-4]]), 
                    raw_elevation=xmlValue(tds[[l-3]], recursive=F),
                    raw_prominance=xmlValue(tds[[l-2]], recursive=F),
                    raw_isolation=xmlValue(tds[[l-1]], recursive=F),
                    raw_coords=xmlValue(tds[[l]]),
                    stringsAsFactors=FALSE)
                } 
            }))
data$name <- sapply(strsplit(data$raw_name, '[\n[]'), function(l){l[[1]]})
two_unit_col_to_numeric <- function(data, raw_colname) {
    as_str <- stringr::str_match(stringr::str_replace_all(data[,raw_colname], ',', ''), '([0-9.]+)\\W([a-zA-Z]+)\\n([0-9.]+)\\W+([a-zA-Z]+)')
    v1 <- as.numeric(as_str[,2])
    u1 <- levels(as.factor(as_str[,3]))
    v2 <- as.numeric(as_str[,4])
    u2 <- levels(as.factor(as_str[,5]))
    stopifnot(length(u1)==1)
    stopifnot(length(u2)==1)
    colname <- stringr::str_replace(raw_colname, '^raw_', '')
    data[,sprintf("%s_%s", colname, u1[[1]])] <- v1
    data[,sprintf("%s_%s", colname, u2[[1]])] <- v2
    return(data)
}
data <- two_unit_col_to_numeric(data, 'raw_elevation')
data <- two_unit_col_to_numeric(data, 'raw_prominance')
data <- two_unit_col_to_numeric(data, 'raw_isolation')
data$lat <- as.numeric(stringr::str_match(data$raw_coords, '/ ([-0-9.]+);')[,2])
data$lng <- as.numeric(stringr::str_match(data$raw_coords, ';\\W([-0-9.]+)')[,2])
```

## Interactive Plot


```{r, fig.height=3.5, fig.width=5}
library(ggplot2, verbose=FALSE, warn.conflicts=FALSE, quietly=TRUE)
library(plotly, verbose=FALSE, warn.conflicts=FALSE, quietly=TRUE)

p <- qplot(prominance_m, isolation_km, data=data)+
    scale_x_log10(breaks=c(250,500,750,1000,1500,2000,4000,6000))+
    scale_y_log10(breaks=c(1, 5, 20, 50, 250, 1000, 2000, 5000))
ggplotly(p)
```


