# EDA: Extended summary of all columns

I like to look at the columns in a competition, with more detail than the R
summary function provides.


## Code

In hope it helps others, here is an R function to provide extra detail.

```{r}
# Summarize a data frame. Shows the (first) class, the number of
# unique levels, the min and max, as well as median and hinges
# (quartiles).  For non-numeric columns (EG, character, factors) the
# mean, median and hinges do not make sense, but this function can and
# will still show min and max.  Beware: this will turn the min/max
# output into strings for the numeric columns too, since R wants
# eveything in a column to have the same type!

df.nlevels <- function(data)
{
    do.call(rbind,lapply(names(data),function(x){
        klass <- class(data[,x])[[1]]
        retval <- data.frame(name=x, klass=klass, nlevels=length(unique(data[,x])), NAs=sum(is.na(data[,x])), mean=NA, min=NA, lower.hinge=NA, median=NA, upper.hinge=NA, max=NA)
        if (klass %in% c('integer', 'numeric','POSIXct')) {
            retval[,'mean'] <- mean(data[,x],na.rm=TRUE)
            retval[,6:10] <- fivenum(data[,x],na.rm=TRUE)
        } else if (klass %in% c('character')) {
            retval[,'min'] <- min(data[,x], na.rm=TRUE)
            retval[,'max'] <- max(data[,x], na.rm=TRUE)
        } else if (klass %in% c('factor','ordered')) {
            lvls <- levels(data[,x])
            retval[,'min'] <- lvls[1]
            retval[,'max'] <- lvls[length(lvls)]
        }
        return(retval)
    }))
}
```


## Results

Of course, a function isn't helpful until you use it to look at the
results. I'm sorry so many columns wrap here. I suggest a spreadsheet
program to view the CSV.


```{r, echo=FALSE}
# Boilerplate load data code

input_dir = '../input'
input_filename <- function(name, dir=input_dir) {
    file.path(input_dir, name)
}
predict_col <- 'price_doc'
macro <- read.csv(input_filename('macro.csv'))
overall_train <- merge(read.csv(input_filename('train.csv')), macro, by='timestamp')

# set seed as contest completion date for repeatability
set.seed(20170529)

# Break overall_train into three sets, train, validate and test
# as 60:20:20 split
library(caret)
train_idx <- caret::createDataPartition(overall_train[,predict_col], p=0.6, list=FALSE)
train <- overall_train[train_idx,]
non_train <- overall_train[-c(train_idx),]
validation_idx <- caret::createDataPartition(non_train[,predict_col], p=0.5, list=FALSE)
validation <- non_train[validation_idx,]
test <- non_train[-c(validation_idx),]
```

```{r}
(train_summary <- df.nlevels(train))
write.csv(train_summary, file='train_summary.csv', row.names=FALSE)
```
