---
title: "School Performance^[Copyright 2016 R. A. Reitmeyer. Released under Creative Commons CC-BY 4.0 license.]"
date: August 2016
output: beamer_presentation
---

## Exercise

* Your objective for the rest of the night is to try to model
  school district performance on the CAASPP for some bay area
  communities
    * CA Assessment of Student Performance and Progress
  
* QUESTION: What is "percentage standard met and above", 
   in terms of other columns?
    - Excluding other percentage..standard columns, of course!
    
* What's the best model for this, in the smallest-BIC sense?

* Data was assembled from CA Department of Education, US Census
  and was turned into geoJSON maps
  
* See http://www.opensmc.org/LD3 for basic visualization

* Please grab data, and these slides, from 
  github://rareitmeyer/linear_regression_class


## Load Into R

* Won't use mapping for this, so use jsonlite package
    * Install if you don't have it via install.packages('jsonlite')
   
* Read in the data and drop the geo bits
```{r}
library(jsonlite)
g5m_geo <- jsonlite::fromJSON(
    'grade_5_mathematics.geojson')
g5m <- g5m_geo$features$properties
dim(g5m)
```

## BIC

* As you try models, check the BIC

```{r}
BIC(lm(percentage_standard_met_and_above ~ 
           total_caaspp_enrollment, data=g5m))
BIC(lm(percentage_standard_met_and_above ~ 
           CurrentExpenseADA, data=g5m))
```

* Advice: Don't hop straight into model-making. Start with graphs!
    * library(ggplot2) and qplot(x, y, data=g5m) are your friends!
    * Use summary(g5m) to see all columns and some info
    * Install the car package and look at residualPlot or residualPlots!



