---
title: 'Data Products Week 2: North American Peaks'
author: "R. A. Reitmeyer"
date: "February 16, 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

This is a simple presentation of North American Mountain Peaks
in a leaflet map. It was created February 16, 2017 for the 
week 2 project of the "Data Products" coursera course.


## Data Source

Data comes from a wikipedia:
[https://en.wikipedia.org/wiki/Table_of_the_highest_major_summits_of_North_America]

A local copy of this page was saved 2017-02-16 to insure consistency.

Using the KML file would have been much easier, but the KML lacks the elevation,
prominance and isolation information.

## Data Processing

Data was extracted from the main table in the HTML page and parsed.

```{r}
library(XML)
library(stringr)

tree <- htmlParse('./en.wikipedia.org_wiki_Table_of_the_highest_major_summits_of_North_America.html')
data <- do.call(rbind, 
    xpathApply(tree, '/html/body//table[@class="wikitable sortable"]/tr',
        function(tr) {
            tds<-getNodeSet(tr, "./td")
            l <- length(tds)
            if (l > 7) { 
                data.frame(
                    raw_name=xmlValue(tds[[l-6]]), 
                    region=xmlValue(tds[[l-5]]), 
                    range=xmlValue(tds[[l-4]]), 
                    raw_elevation=xmlValue(tds[[l-3]], recursive=F),
                    raw_prominance=xmlValue(tds[[l-2]], recursive=F),
                    raw_isolation=xmlValue(tds[[l-1]], recursive=F),
                    raw_coords=xmlValue(tds[[l]]),
                    stringsAsFactors=FALSE)
                } 
            }))
data$name <- sapply(strsplit(data$raw_name, '[\n[]'), function(l){l[[1]]})
two_unit_col_to_numeric <- function(data, raw_colname) {
    as_str <- stringr::str_match(stringr::str_replace_all(data[,raw_colname], ',', ''), '([0-9.]+)\\W([a-zA-Z]+)\\n([0-9.]+)\\W+([a-zA-Z]+)')
    v1 <- as.numeric(as_str[,2])
    u1 <- levels(as.factor(as_str[,3]))
    v2 <- as.numeric(as_str[,4])
    u2 <- levels(as.factor(as_str[,5]))
    stopifnot(length(u1)==1)
    stopifnot(length(u2)==1)
    colname <- stringr::str_replace(raw_colname, '^raw_', '')
    data[,sprintf("%s_%s", colname, u1[[1]])] <- v1
    data[,sprintf("%s_%s", colname, u2[[1]])] <- v2
    return(data)
}
data <- two_unit_col_to_numeric(data, 'raw_elevation')
data <- two_unit_col_to_numeric(data, 'raw_prominance')
data <- two_unit_col_to_numeric(data, 'raw_isolation')
data$lat <- as.numeric(stringr::str_match(data$raw_coords, '/ ([-0-9.]+);')[,2])
data$lng <- as.numeric(stringr::str_match(data$raw_coords, ';\\W([-0-9.]+)')[,2])
```

## Map 1: Peaks, showing isolation

Here are the peaks, with circles representing isolation (distance to nearest 
higher peak).  Ideally, the circle for each peak is exactly big enough to just
touch the nearest peak with higher elevation. In practice, the Mercator 
projection used by default does not preserve distances, especially at continent scale.


```{r, fig.height=4, fig.width=6}
library(leaflet)

map <- leaflet() %>%
    addTiles() %>%
    addCircles(lat=data$lat, lng=data$lng, weight=1, radius=data$isolation_km*1000) %>%
    addMarkers(lat=data$lat, lng=data$lng, popup=sprintf("%s, %0.0f m", data$name, data$elevation_m))
map
```


## Replacing Mercator, step 1:

It should be possible to ask modern version of leaflet to reproject to
an azimuthal equaldistant projection, which would preserve distances. 
Something like "+proj=aeqd +lat_0=45 +lon_0=-105" should be fine.
(See [https://bhaskarvk.github.io/leaflet/projections.html].)

However, leaflet will expect tiles to be in the new projection, and
there are no azimuthal equadistant tiles handy. 

Without an underlying map it is difficult to see where things are, however.

So the next map uses a vector world map available from [https://d3js.org/world-110m.v1.json], 
as a TopoJSON file. (dowloaded 2017-02-16.) 

## Map 2: No tiles

This is still in Mercator projection, but tiles have been replaced
by a vector world map. Note that leaflet does not handle
lines that cross +/- 180 degrees of longitude correctly.

```{r, fig.height=4, fig.width=6}
fp <- file('world-110m.v1.json')
tj <- readLines(fp)
close(fp)

map2 <- leaflet() %>%
    addTopoJSON(tj, layerId='land', color='black', weight=1) %>%
    addCircles(lat=data$lat, lng=data$lng, weight=1, radius=data$isolation_km*1000) %>%
    addMarkers(lat=data$lat, lng=data$lng, popup=sprintf("%s, %0.0f m", data$name, data$elevation_m))
map2
```

## Map 3

There is no map 3, unfortuntely.

Reprojection requires 1.1, which is not available yet. The R repos
still have 1.0.1: [https://cran.r-project.org/web/packages/leaflet/index.html]



