---
title: "Wind and Flood Most Damaging Categories of Severe Weather in United States from 1950-2011"
author: "R. A. Reitmeyer"
date: "October 24, 2016"
output: pdf_document
---

# Synopsis

This paper examines the impact of severe weather on the health and
economy of the United States in the period from 1950 through 2011,
as motivated by the Coursera "Reproduceable Research" class. NOAA
data on weather have been downloaded and processed in R to determine
the most severe category of weather, in terms of deaths and damages.
To simplify for the civic planner, weather has been categorized as
cold, drought, flood, heat, precipitation, wind and other. Raw data
has been processed to calculate damage dollar values in adjusted
year 2000 dollars, and aggregated to compute totals per category. 
Wind-related events have caused the most deaths, while flood events 
have caused the most damages. Civic planners should invest in protecting
against these hazards.


# Data Processing

## Origin

All data in this paper originates from NOAA, via a copy stored for this 
course on cloudfront.net.

```{r load_raw, cache=TRUE}
data_filename <- 'repdata_data_StormData.csv.bz2'
if (!file.exists(data_filename)) {
    download.file(
        'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2', 
        data_filename, method='curl')
}
raw_data <- read.csv(bzfile(data_filename))
```


## Data Transformation

The raw data has been transformed by catgorizing event types, processing damage
dollar values, and aggregating into totals-by-category.


### Event Categories

```{r summarize_raw_evtype_levels}
# Would inline this, but course instuctions emphasize showing all the analysis via
# grading criteria #13, "Do all the results of the analysis (i.e. figures, tables, 
# numerical summaries) appear to be reproducible?"
raw_evtype_levels <- nlevels(raw_data$EVTYPE)
explicit_wind_evtype_levels <- sum(grepl('wind', levels(raw_data$EVTYPE), ignore.case=TRUE))
example_odd_evtype_labels <- paste(grep(
    'apache|driest|no severe|northern lights|flag criteria|southeast',
    levels(raw_data$EVTYPE), ignore.case=TRUE, value=TRUE), collapse='", "')
```
The raw data contains `r raw_evtype_levels` distinct types of 
events. Many of these are variations of a single underlying 
weather pattern. Others appear to be quite rare phenonoma or 
data-entry issues.

* 'Wind' explicitly appears in `r explicit_wind_evtype_levels` 
    of the names, without counting implicit names like "tornado".
* Event types include odd values: "`r example_odd_evtype_labels`", ...

Since no civic planner should worry about the distinction between
`r explicit_wind_evtype_levels` kinds of wind, the raw data event 
types have been aggregated into a much smaller number of categories
that civic planners can protect against. An "other" bucket is included
so the reader can confirm that the top categories for death and damage
have not been accidentally omitted.

* Cold
* Drought
* Flood
* Heat
* Precipitation (rain/snow)
* Wind
* Other

For example, a civic planner can stockpile emergency blankets and
heating oil for cold emergencies, and encourage more insulation
as part of preparedness for cold.

Note that a real-world weather system can fall into more than one of 
these categories. In this case, for this simplified analysis, no
effort has been made to refine the relative proportions for 
each simplified category. For example, a hurricane produces water
damage as well as wind damage, and this analysis simply assigns 
1/2 of the hurricane's destruction to each category. 

For purposes of the analysis, events categorized as cold-related are 
those with event type names matching "blizzard," "chill," "cold," 
"freeze," "frost," ice," hyopthermia", "low temp," and "snow."

Event types categorized as drought are those matching 
"below normal precip," "drought," "driest," "dry," "lack of snow," 
and "low rain."

Flood-related events match "flood," "surge" and "tide."

Heat-related events match "heat," "high temp," "hot," "record high," and "warm."

Precipitation-related events match "blizzard," "tropical depression," "hail," "hurricane," "precip," "rain," "shower," "snow," "storm," "tstm," and "wet".

Wind-related events match "downburst," "dust," "funnel," "hurricane," "microburst," "tornado," "tropical depression," "tropical storm," "typhoon," "waterspout," "whirlwind," and "wind."


```{r event_cat}
cat_terms <- data.frame(EVTYPE=as.character(levels(raw_data$EVTYPE)),
                    cold=0,drought=0,flood=0,heat=0,precipitation=0,wind=0,other=0)
cold_idx = grep(
    'blizzard|chill|cold|freeze|frost|ice|hypothermia|low temp|snow',
    cat_terms$EVTYPE, ignore.case=TRUE)
cat_terms[cold_idx,'cold'] <- 1
drought_idx = grep(
    'below normal precip|drought|driest|dry|lack of snow|low rain',
    cat_terms$EVTYPE, ignore.case=TRUE)
cat_terms[drought_idx,'drought'] <- 1
flood_idx <- grep(
    'flood|surge|tide',
    cat_terms$EVTYPE, ignore.case=TRUE)
cat_terms[flood_idx,'flood'] <- 1
heat_idx <- grep(
    'heat|high temp|hot|record high|warm',
    cat_terms$EVTYPE, ignore.case=TRUE)
cat_terms[heat_idx,'heat'] <- 1
precip_idx <- grep(
    'blizzard|tropical depression|hail|hurricane|precip|rain|shower|snow|storm|tstm|wet',
    cat_terms$EVTYPE, ignore.case=TRUE)
cat_terms[precip_idx,'precipitation'] <- 1
wind_idx <- grep(
    'downburst|dust|funnel|hurricane|microburst|tornado|tropical depression|tropical storm|typhoon|waterspout|whirlwind|wind',
    cat_terms$EVTYPE, ignore.case=TRUE)
cat_terms[wind_idx,'wind'] <- 1

# set 'other' if nothing else is selected for a row.
term_sum <- rowSums(cat_terms[,-1])
cat_terms[term_sum == 0,'other'] <- 1

# For each row, divide by the sum to gets weights to sum to 1.
term_sum <- rowSums(cat_terms[,-1])
for (col in 2:ncol(cat_terms)) {
    cat_terms[,col] <- cat_terms[,col]/term_sum
}

# Merge categories with raw data
data <- merge(raw_data, cat_terms, by='EVTYPE')

# Save off the column numbers for the weights, which run from 
# 'cold' to 'other.'
cold_col <- which(names(data)=='cold')
other_col <- which(names(data)=='other')

# Sanity check: every row's category matrix columns should 
# sum to 1
stopifnot(rowSums(data[,cold_col:other_col])==1)
```


### Handling Damage Dollar Values

The raw data encodes the dollar value of property and crop
damages by separately storing a mantissa-like three-digit value 
and a separate 'exponent'. Exponents are intended to be
'K' for thousands, 'M' for millions, and 'B' for (US) billions,
(and blank for 'ones').

However, some other exponents have crept in, suggesting data
errors.

For purpose of this analysis, damages associated with unrecognized
exponents are treated as missing (NA) values and ignored.

Since the data spans many years, inflation cannot be ignored: a dollar
in 2011 was worth about 11 cents of what a dollar was worth in 1950
[http://www.bls.gov/data/inflation_calculator.htm]. To account for
this, all dollar figures are reported in adjusted year-2000 dollars
or billions of year-2000 dollars.

```{r handle_damage_dollar_values}
# use the ACS package to handle inflation
library(acs, quietly=TRUE, verbose=FALSE, warn.conflicts=FALSE)  
data(cpi)                   # get CPI data, a vector of values relative
                            # to 1982-1984, with names matching the year.
library(lubridate, quietly=TRUE, verbose=FALSE, warn.conflicts=FALSE)

# Utility function to convert damage data into numbers. Raw data uses
# exponents 'K' for thousands, 'M' for millions and 'B' for billions;
# some raw data is encoded with other exponents (EG, '?', '-', '8')
# which suggest data entry errors. Data for all such 'ambiguous' 
# exponents are returned as NA values.
dmg2num <- function(data, dmgcol) {
    expabbr <- toupper(data[,paste(dmgcol,'EXP',sep='')])
    expval <- ifelse(expabbr == '' | expabbr == ' ', 1,
                     ifelse(expabbr == 'K', 1000,
                            ifelse(expabbr == 'M', 1000^2,
                                   ifelse(expabbr == 'B', 1000^3,
                                          NA))))
    data[,dmgcol] * expval
}
# calculate damages, in nominal dollars, and save which damages are not-NA.
data$propdmg_num <- dmg2num(data, 'PROPDMG')
data$cropdmg_num <- dmg2num(data, 'CROPDMG')
data$dmg_num <- with(data, propdmg_num+cropdmg_num)
data$dmg_nonna <- !is.na(data$dmg_num)

# get year and CPI for the year.
data$year <- lubridate::year(as.Date(as.character(data$BGN_DATE),
                                     '%m/%d/%Y %H:%M:%S'))
# confirm fast (index-based) CPI lookup gets the right years.
stopifnot(
    as.numeric(names(cpi)[data$year-min(as.numeric(names(cpi)))+1]) == 
    data$year)
data$cpi <- cpi[data$year-min(as.numeric(names(cpi)))+1]

# save off the year 2000 CPI to normalize with
cpi_2000 <- cpi[2000-min(as.numeric(names(cpi)))+1]
```

### Aggregation

Within the analysis, each event type is represented by a row 
in a event category weighting matrix. For example, a row for
the 'hurricane' event type has 0 for cold, drought, flood, and
heat columns, but has values 0.5 for precipitation and wind 
columns.

Deaths and damages have been aggegated to categories via
simple matrix algebra, multipling fatalities and damages
after handing any NA rows.

As described above, inflation is handled by using the consumer
price index to adjust damage dollar figures to use year-2000 dollars.

```{r aggregate_event_categories}
# Matrix math: 
# * Totals for fatalites per category is simply 
#   fatalties * category_weights.
# * Totals for damages are more complex. 
#   We must ignore the NA rows by subsetting on
#   damages-are-not-NA, and then we need to deal
#   with the CPI by multiplying by a correction of
#   CPI[2000]/CPI[year]. To make scales human-readable
#   compute damages in (US) Billions.
totals <- rbind(
    t(data[,'FATALITIES'])%*%as.matrix(data[,cold_col:other_col]),
    t(data[data$dmg_nonna,'dmg_num']*cpi_2000/data[data$dmg_nonna,'cpi']/1e9)%*% 
        as.matrix(data[data$dmg_nonna,cold_col:other_col])
)
row.names(totals) <- c('deaths','damages_2000_Bdollars')

# Not worth requireing reshape to convert to long; there's just
# two rows for the data of interest. So write a little function
# to convert a two row matrix to a "long" data frame.
two_row_matrix_to_long <- function(data) {
    x <- c(data[1,],data[2,])
    retval <- data.frame(
        category=names(x), 
        var=row.names(data)[rep(1:2,each=ncol(data))],
        value=x)
}
totals_long <- two_row_matrix_to_long(totals)
```


\newpage

# Results

Civic planners should prepare for wind events, the leading cause of
deaths and second most prominent cause of damages. Flood is the
leading cause of damage, but relatively small source of fatalities.

## Events Most Damaging to Public Health (Fatalities)

```{r}
# confirm totals agree and calculate total deaths
stopifnot(sum(raw_data$FATALITIES) == sum(subset(totals_long,var=='deaths')$value))
total_deaths = sum(raw_data$FATALITIES)
```

```{r deaths_chart, fig.width=6,fig.height=3,fig.cap='Bar chart of total death counts by severe weather category, United States, 1950-2011.'}
library(ggplot2)
ggplot(subset(totals_long, var=='deaths'), aes(x=category, y=value))+
    geom_bar(stat='identity')+
    ylab('Deaths')+
    ggtitle('Deaths by Category of Severe Weather')
```

During the 1950-2011 period, there were `r sprintf('%0.0f',total_deaths)`
deaths related to severe weather events in the US.

Wind events (which include tornados and hurricanes) cause the most
fatalities. Civic planners in the US would be well advised to 
investigate methods to deal with high winds.

Heat-related events are the second-highest category of fatality.

"Other," a catch-all including lighting, fire, "Apache County"
and other miscellaneous kinds of events, is the third most serious
category of fatalities. Other is less than 1/3 of the deaths of 
wind events, and roughly 1/5th the deaths from the top-two categories.


## Events Most Damaging to the Economy

```{r}
# confirm totals agree (within $1) after CPI adjustment, 
# and calculate total damages
stopifnot(abs(
        sum(data$dmg_num*cpi_2000/data$cpi, na.rm=TRUE) -
        sum(subset(totals_long,var=='damages_2000_Bdollars')$value*1e9))
    <= 1)
total_damages_2000_Bdollars = sum(data$dmg_num*cpi_2000/data$cpi/1e9, na.rm=TRUE)
```

```{r damages_chart, fig.width=6,fig.height=3,fig.cap='Bar chart of total damages by severe weather category, United States, 1950-2011, adjusted year-2000 dollars.'}
library(ggplot2)
ggplot(subset(totals_long, var=='damages_2000_Bdollars'), aes(x=category, y=value))+
    geom_bar(stat='identity')+
    ylab('Damages, Billions of Dollars')+
    ggtitle('Damages by Category of Severe Weather')
```

During the 1950-2011 period, severe weather events
cost the United States at least `r sprintf('%0.0f',total_damages_2000_Bdollars)`
billions of dollars in damages (adjusted for inflation to 
year-2000 dollars). Note that this is a minimum; as discussed
above, suspect damage data has been ignored.

Flooding is the most damaging category of event, followed by
wind and precipitation. This is in line with expectation:
recent hurricanes like Katrina and Sandy have caused incredible
amounts of damage from wind, rain and associated flooding, with
flooding the most-maligned source of damage.


\newpage

# Appendix: Reproduceability Notes

To reproduce this paper, you will need a copy of:

* The Rmarkdown file for this paper
* R 3.2 or later
* Rstudio
* R packages
    - acs (for inflation from 1950 to today, as consumer price index)
    - lubridate
    - ggplot2
* Web access to download the original NOAA data from the cloudfront site



